---
title: "GBM single-cell"
author: "Maksym"
date: "7/21/2022"
output: html_document
---

Packages
```{r}
suppressPackageStartupMessages({
  library(Seurat)
  library(stringr)
  library(purrr)
  library(tidyverse)
  library(DoubletFinder)
  library(RColorBrewer)
})
```

Parallel execution
```{r}
library(doParallel)
library(foreach)
no_cores <- detectCores()
cl <- makeCluster(no_cores)
registerDoParallel(cl)
```

Adds necessary metadata for sclean algo
```{r, echo=F}
add_meta_seurat <- function(obj) {
  obj$log10GenesPerUMI <- log10(obj$nFeature_RNA) / log10(obj$nCount_RNA)
  
  pct_counts_in_top_20_features_fun <- function(cell, obj){
    sum(sort(obj@assays$RNA@counts[,cell], decreasing = TRUE)[1:20])/sum(obj@assays$RNA@counts[,cell])
  }
  
  x = foreach(cell=Cells(obj), .packages = "Seurat") %dopar% pct_counts_in_top_20_features_fun(cell, obj)
  obj$pct_counts_in_top_20_features = unlist(x)
  
  obj$mitoRatio <- PercentageFeatureSet(object = obj, pattern = "^MT-")
  obj$mitoRatio <- obj@meta.data$mitoRatio / 100
  obj$rpsRatio <- PercentageFeatureSet(object = obj, pattern = "^RPS-")
  obj$rplRatio <- PercentageFeatureSet(object = obj, pattern = "^RPL-")
  obj$riboRatio <- obj$rpsRatio + obj$rplRatio
  obj@meta.data <- obj@meta.data %>%
    dplyr::rename(seq_folder = orig.ident,
                  nUMI = nCount_RNA,
                  nGene = nFeature_RNA)
  return(obj)
}

filter_seurat_paper <- function(obj) {
  subset(obj, subset = nGene >= 500 & mitoRatio <= .2 & riboRatio <= .5)
}
```

DoubetFinder function for Seurat
```{r}
detect_doublets <- function(obj) {
  obj <- NormalizeData(object = obj)
  obj <- FindVariableFeatures(object = obj)
  obj <- ScaleData(object = obj)
  obj <- RunPCA(object = obj)
  obj <- FindNeighbors(object = obj, dims = 1:10)
  obj <- FindClusters(object = obj)
  obj <- RunUMAP(object = obj, dims = 1:10)
  
  ## pK Identification (no ground-truth) ---------------------------------------------------------------------------------------
  sweep.res.list_pbmc <- paramSweep_v3(obj, PCs = 1:10, sct = FALSE)
  sweep.stats_pbmc <- summarizeSweep(sweep.res.list_pbmc, GT = FALSE)
  bcmvn_pbmc <- find.pK(sweep.stats_pbmc)
  
  pK <- bcmvn_pbmc %>%
    filter(BCmetric == max(BCmetric)) %>%
    pull(pK) %>%
    as.numeric()
  
  ## Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
  annotations <- obj@meta.data$seurat_clusters
  homotypic.prop <- modelHomotypic(annotations)
  nExp_poi <- round(0.076*nrow(obj@meta.data))
  nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
  
  # run doubletFinder 
  obj <- doubletFinder_v3(obj, 
                                       PCs = 1:10, 
                                       pN = 0.25, 
                                       pK = pK, 
                                       nExp = nExp_poi.adj,
                                       reuse.pANN = FALSE,
                                       sct = FALSE)
  
  obj$DoubletFinder = obj@meta.data %>%
    select(starts_with("DF.")) %>%
    pull()
  
  obj@meta.data = obj@meta.data %>%
    rownames_to_column() %>%
    select(-c(starts_with("DF."), starts_with("pANN"))) %>%
    column_to_rownames()
  
  return(obj)
}
```

## Reading data

```{r}
doublet_dir = "/Volumes/TOSHIBA/TIMElab/GSE182109_processed/scrublet/"
data_dir = "/Volumes/TOSHIBA/TIMElab/GSE182109_RAW/"
sample_ids = list.files(data_dir)
sample_ids
```
Reading 10x data
```{r,eval=F}
read_10X_into_list <- function(file) {
  doublets <- readLines(paste0(doublet_dir, file, ".txt"))
  seurat_data <- Read10X(data.dir = paste(data_dir, file,sep="/"))
  seurat_obj <- CreateSeuratObject(counts = seurat_data,
                                   project = str_split(file, "_", Inf, T)[,1])
  seurat_obj$scrublet.doublet <- doublets
  #seurat_obj <- add_meta_seurat(seurat_obj)
  return(seurat_obj)
}

obj_list = lapply(sample_ids, read_10X_into_list)
```

```{r}

meta_list = lapply(obj_list, function(x) {
  x@meta.data %>% rownames_to_column("cell")
})

meta = bind_rows(meta_list, .id = "id")

meta = meta %>% unite("cell", c("cell", "id"), sep="_")
```



Saving the unedited object
```{r,eval=F}
saveRDS(obj_list, "E:/TIMElab/GSE182109_processed/obj_list.rds")
```


## Filtering

I ran sclean on the obj_list. From the diagnostic images, samples #9 and #10 had too few cells in them.
Removing those samples.
```{r}
obj_list_clean = lapply(obj_list_clean, function(x) {x[,x$qc_default==1]})

# removing suspicious samples
obj_list_clean[[9]] <- NULL
obj_list_clean[[10]] <- NULL
```

```{r, eval=F}
#obj_list = lapply(obj_list, detect_doublets)
```

I ended up not running DoubletFinder because the result looked odd - the "doublets" were distributed pretty much randomly over the embedding. Will have to find out what's the problem.

```{r,eval=F}
saveRDS(obj_list, "E:/TIMElab/GSE182109_processed/obj_list.rds")
```

rPCA Integration script was run on the samples.


Reading the integrated object ...
```{r}
DefaultAssay(seurat_obj) <- "integrated"

# Run the standard workflow for visualization and clustering
seurat_obj <- ScaleData(seurat_obj, verbose = FALSE)
seurat_obj <- RunPCA(seurat_obj, npcs = 30, verbose = FALSE)
seurat_obj <- RunUMAP(seurat_obj, reduction = "pca", dims = 1:30)
seurat_obj <- FindNeighbors(seurat_obj, reduction = "pca", dims = 1:30)
seurat_obj <- FindClusters(seurat_obj, resolution = 0.6)
Idents(seurat_obj) <- "integrated_snn_res.0.6"
```

```{r}
DimPlot(seurat_obj, label = T)
```


```{r}
freq_df = seurat_obj@meta.data %>%
  group_by(seq_folder,integrated_snn_res.0.6) %>%
  summarise(Count = n()) %>%
  mutate(Prop = Count/sum(Count))

ggplot(data = seurat_obj@meta.data, aes(x = seq_folder, fill = integrated_snn_res.0.6)) + 
  geom_histogram(colour = 'white', position = 'fill', stat = "count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
freq_df = seurat_obj@meta.data %>%
  group_by(integrated_snn_res.0.6, seq_folder) %>%
  summarise(Count = n()) %>%
  mutate(Prop = Count/sum(Count))

ggplot(data = seurat_obj@meta.data, aes(x = integrated_snn_res.0.6, fill = seq_folder)) + 
  geom_histogram(colour = 'white', position = 'fill', stat = "count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
seurat_obj$lognUMI <- log2(seurat_obj$nUMI)
VlnPlot(seurat_obj, pt.size = 0, "lognUMI") / VlnPlot(seurat_obj,  pt.size = 0,  "mitoRatio")
```


Removing GSM5518629 because it's mostlty cluster number 3 which I am also removing
Clusters 17 and 19 is probably apoptotic
```{r}
seurat_obj = seurat_obj[,seurat_obj$seq_folder!="GSM5518629"]
seurat_obj = seurat_obj[,seurat_obj$integrated_snn_res.0.6!=5]
seurat_obj = seurat_obj[,seurat_obj$integrated_snn_res.0.6!=17]
seurat_obj = seurat_obj[,seurat_obj$integrated_snn_res.0.6!=19]
```

### Cluster 11
```{r}
seurat_obj_sub = subset(seurat_obj, idents = 11)
VlnPlot(seurat_obj_sub, pt.size = 0, "lognUMI") +
  geom_hline(yintercept = 10.5)
```
```{r}
cells_to_remove = colnames(seurat_obj_sub)[seurat_obj_sub$lognUMI<10.5]
```


### Cluster 7
```{r}
seurat_obj_sub = subset(seurat_obj, idents = 7)
VlnPlot(seurat_obj_sub, pt.size = 0, "mitoRatio") +
  geom_hline(yintercept = 0.025)
```

```{r}
cells_to_remove = c(cells_to_remove, colnames(seurat_obj_sub)[seurat_obj_sub$mitoRatio<.025])
```



### Cluster 21
```{r}
seurat_obj_sub = subset(seurat_obj, idents = 21)
VlnPlot(seurat_obj_sub, pt.size = 0, "lognUMI") +
  geom_hline(yintercept = 11)
```

```{r}
cells_to_remove = c(cells_to_remove, colnames(seurat_obj_sub)[seurat_obj_sub$lognUMI<11])
```


Cluster 22 looks strange but I will recluster and then check if its maps to a cell type

### Cluster 23
```{r}
seurat_obj_sub = subset(seurat_obj, idents = 23)
VlnPlot(seurat_obj_sub, pt.size = 0, "lognUMI") +
  geom_hline(yintercept = 11)
```

```{r}
cells_to_remove = c(cells_to_remove, colnames(seurat_obj_sub)[seurat_obj_sub$lognUMI<11])
```

Removing `r length(cells_to_remove)` cells

```{r}
seurat_obj = seurat_obj[,!colnames(seurat_obj) %in% cells_to_remove]
```

Reclustering
```{r}
# Run the standard workflow for visualization and clustering
seurat_obj <- ScaleData(seurat_obj, verbose = FALSE)
seurat_obj <- RunPCA(seurat_obj, npcs = 30, verbose = FALSE)
seurat_obj <- RunUMAP(seurat_obj, reduction = "pca", dims = 1:30)
seurat_obj <- FindNeighbors(seurat_obj, reduction = "pca", dims = 1:30)
seurat_obj <- FindClusters(seurat_obj, resolution = 0.3)
seurat_obj <- FindClusters(seurat_obj, resolution = 0.6)
Idents(seurat_obj) <- "integrated_snn_res.0.3"
```

```{r}
DimPlot(seurat_obj, label = T)
```

```{r}
VlnPlot(seurat_obj, pt.size = 0, "lognUMI") / VlnPlot(seurat_obj,  pt.size = 0,  "mitoRatio")
```

```{r}
freq_df = seurat_obj@meta.data %>%
  group_by(integrated_snn_res.0.3, seq_folder) %>%
  summarise(Count = n()) %>%
  mutate(Prop = Count/sum(Count))

ggplot(data = seurat_obj@meta.data, aes(x = integrated_snn_res.0.3, fill = seq_folder)) + 
  geom_histogram(colour = 'white', position = 'fill', stat = "count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
freq_df = seurat_obj@meta.data %>%
  group_by(seq_folder,integrated_snn_res.0.3) %>%
  summarise(Count = n()) %>%
  mutate(Prop = Count/sum(Count))

ggplot(data = seurat_obj@meta.data, aes(x = seq_folder, fill = integrated_snn_res.0.3)) + 
  geom_histogram(colour = 'white', position = 'fill', stat = "count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Removing cluster 10
Removing cluster 16 because 50 perc of cells map to only one sample
Filtering lower UMI distributon of 3rd cluster

```{r}
seurat_obj = seurat_obj[,seurat_obj$integrated_snn_res.0.3!=10]
seurat_obj = seurat_obj[,seurat_obj$integrated_snn_res.0.3!=16]
```

### Cluster 3
```{r}
seurat_obj_sub = subset(seurat_obj, idents = 3)
VlnPlot(seurat_obj_sub, pt.size = 0, "lognUMI") +
  geom_hline(yintercept = 10.75)
```

```{r}
cells_to_remove = colnames(seurat_obj_sub)[seurat_obj_sub$lognUMI<10.75]
length(cells_to_remove)
seurat_obj = seurat_obj[,!colnames(seurat_obj) %in% cells_to_remove]
```

```{r}
# Run the standard workflow for visualization and clustering
seurat_obj <- ScaleData(seurat_obj, verbose = FALSE)
seurat_obj <- RunPCA(seurat_obj, npcs = 30, verbose = FALSE)
seurat_obj <- RunUMAP(seurat_obj, reduction = "pca", dims = 1:12)
seurat_obj <- FindNeighbors(seurat_obj, reduction = "pca", dims = 1:12)
seurat_obj <- FindClusters(seurat_obj, resolution = 0.3)
Idents(seurat_obj) <- "integrated_snn_res.0.3"
```
```{r}
DimPlot(seurat_obj, label = T)
```
```{r}
VlnPlot(seurat_obj, pt.size = 0, "lognUMI") / VlnPlot(seurat_obj,  pt.size = 0,  "mitoRatio")
```


```{r}
seurat_obj = seurat_obj[,seurat_obj$integrated_snn_res.0.3!=11]
```

Re-integrating the object ...

```{r}
DimPlot(seurat_obj, label = T, group.by = "integrated_snn_res.0.2")
```

```{r}
freq_df = seurat_obj@meta.data %>%
  group_by(integrated_snn_res.0.2, seq_folder) %>%
  summarise(Count = n()) %>%
  mutate(Prop = Count/sum(Count))

ggplot(data = seurat_obj@meta.data, aes(x = integrated_snn_res.0.2, fill = seq_folder)) + 
  geom_histogram(colour = 'white', position = 'fill', stat = "count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



```{r}
saveRDS(seurat_obj, "E:/TIMElab/GSE182109_processed/seurat_integrated_filtered.rds")
```

```{r}
seurat_obj = readRDS("H:/TIMElab/GSE182109_processed/seurat_integrated_filtered.rds")
```




Let's try running DoubletFinder
```{r,eval=F}
DefaultAssay(seurat_obj) = "RNA"
seurat_obj_list = SplitObject(seurat_obj, split.by = "seq_folder")
seurat_obj_list_test = lapply(seurat_obj_list[1:7], detect_doublets)
```

```{r,eval=F}
p1 = DimPlot(seurat_obj_list_test[[1]], group.by = "DoubletFinder")
p2 = DimPlot(seurat_obj_list_test[[2]], group.by = "DoubletFinder")
p3 = DimPlot(seurat_obj_list_test[[3]], group.by = "DoubletFinder")
p4 = DimPlot(seurat_obj_list_test[[4]], group.by = "DoubletFinder")
p5 = DimPlot(seurat_obj_list_test[[5]], group.by = "DoubletFinder")
p6 = DimPlot(seurat_obj_list_test[[6]], group.by = "DoubletFinder")
p1 + p2 + p3 + p4 + p5 + p6
```



## Annotation

```{r}
DefaultAssay(seurat_obj) <- "RNA"
seurat_obj = NormalizeData(seurat_obj)
seurat_obj = ScaleData(seurat_obj)
p = DotPlot(seurat_obj,
        features = c("PDGFRA", "SOX2", "OLIG1", "NES", "GFAP", "S100B", "EGFR", "MBP", "PDGFRB", "ACTA2", "PROM1", "PECAM1", "TEK", "P2RY12", "PTPRC", "ITGAM", "ITGAX", "S100A8", "CD68", "MRC1", "NKG7", "CD3E", "CD8A", "CD4", "CD79A", "MKI67"),
        cluster.idents = T) +
 theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p
```
```{r}
ggsave(p, path="H:/TIMElab/figures", filename= "markers_all.pdf", width=8.5, height=3, dpi=700)
```


```{r}
Idents(seurat_obj) <- "integrated_snn_res.0.2"
seurat_obj <- RenameIdents(object = seurat_obj, 
                              "0" = "Myeloid",
                              "1" = "Myeloid",
                              "2" = "Glioma",
                              "3" = "Glioma",
                              "4" = "T cell",
                              "5" = "Glioma",
                              "6" = "Glioma",
                              "7" = "Myeloid",
                              "8" = "Glioma",
                              "9" = "Myeloid",
                              "10" = "Oligo",
                           "11" = "Pericyte",
                           "12" = "B cell",
                           "13" = "Endothelial",
                           "14" = "T cell")
seurat_obj$cell.type = Idents(seurat_obj)

p = DimPlot(seurat_obj, label = TRUE, group.by = "cell.type")
p
```

```{r,eval=F}
saveRDS(seurat_obj, "H:/TIMElab/GSE182109_processed/seurat_integrated_filtered.rds")
```

                                                      
